Parallel Algorithm for Solving Multicriterial Optimization Problems Using Elements of Machine Learning

### 2025-09-19 2025/RSCD_MCO/RSD-2025.tex
fix RSCD editor comments

\begin{document}
%
\title{Parallel algorithm for solving multicriterial optimization problems using elements of machine learning}
%
\titlerunning{Parallel algorithm for solving multicriterial optimization problems}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{{Sergey Konnov\orcidID{0009-0003-4590-0870} \and
Evgeny Kozinov \Letter\orcidID{0000-0001-6776-0096} \and
Konstantin Barkalov \orcidID{0000-0001-5273-2471} \and
Alexander Sysoyev \orcidID{0000-0003-1542-7624} \and
Vladimir Grishagin\orcidID{0000-0002-2884-3670}}
}
%
\authorrunning{S. Konnov et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia 
\email{\{konstantin.barkalov,evgeny.kozinov\}@itmm.unn.ru}, \email{vagris@unn.ru}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Solving problems of multi-criteria optimization involves finding many combinations of optimization parameters, in which the values of the criteria cannot be improved in several criteria at once without worsening the values   of the remaining ones (Pareto set). In the framework of the research, several assumptions were introduced: criteria are multiextremal and difficult to calculate, presented in the form of a ``black box'', and the number of optimized parameters is small. The paper presents a method for solving this class of problems based on the information-statistical approach and a machine learning procedure used to increase search efficiency. A parallel implementation of the method is described, which is supplemented with the ability to perform calculations in asynchronous parallel mode. The efficiency and scalability of the proposed algorithm is analyzed on the base of solving a test class of multiextremal problems.

@@

\section{Problem statement}
\label{sec2}

@@

\section{Computational scheme of the developed algorithm}
\label{sec3}

@@

\subsection{Block of the global search algorithm}
\label{subsec31}

@@

\subsection{Block for accumulating and reusing the information about current state of the search}
\label{subsec32}

@@

\subsection{Block for building the Pareto set estimation}
\label{subsec33}

@@

\subsection{Block of  building the machine learning for estimation of the Pareto set}
\label{subsec34}

@@

\subsection{Block for the choice of trial points and implementation of computational experiments}
\label{subsec35}

@@

\section{Results of computational experiments}
\label{sec4}

>>>>

\begin{document}
%
\title{Parallel Algorithm for Solving Multicriterial Optimization Problems Using Elements of Machine Learning}
%
\titlerunning{  %Parallel Algorithm for Solving Multicriterial Optimization Problems
}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{{Sergey Konnov\orcidID{0009-0003-4590-0870} \and
Evgeny Kozinov \Letter\orcidID{0000-0001-6776-0096} \and
Konstantin Barkalov \orcidID{0000-0001-5273-2471} \and
Alexander Sysoyev \orcidID{0000-0003-1542-7624} \and
Vladimir Grishagin\orcidID{0000-0002-2884-3670}}
}
%
\authorrunning{
%S. Konnov et al.
}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia 
\email{\{konstantin.barkalov,evgeny.kozinov\}@itmm.unn.ru}, \email{vagris@unn.ru}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Solving problems of multi-criteria optimization involves finding many combinations of optimization parameters, in which the values of the criteria cannot be improved in several criteria at once without worsening the values   of the remaining ones (Pareto set). In the framework of the research, several assumptions were introduced: criteria are multiextremal and difficult to calculate, presented in the form of a ``black box'', and the number of optimized parameters is small. The paper presents a method for solving this class of problems based on the information-statistical approach and a machine learning procedure used to increase search efficiency. A parallel implementation of the method is described, which is supplemented with the ability to perform calculations in asynchronous parallel mode. The efficiency and scalability of the proposed algorithm is analyzed on the base of solving a test class of multiextremal problems.

@@

\section{Problem Statement}
\label{sec2}

@@

\section{Computational Scheme of the Developed Algorithm}
\label{sec3}

@@

\subsection{Block of the Global Search Algorithm}
\label{subsec31}

@@

\subsection{Block for Accumulating and Reusing the Information About Current State of the Search}
\label{subsec32}

@@

\subsection{Block for Building the Pareto Set Estimation}
\label{subsec33}

@@

\subsection{Block of Building the Machine Learning for Estimation of the Pareto Set}
\label{subsec34}

@@

\subsection{Block for the Choice of Trial Points and Implementation of Computational Experiments}
\label{subsec35}

@@

\section{Results of Computational Experiments}
\label{sec4}
